### 2013-12-16 paper/paper.tex
Cleaned up paper directory

\documentclass[conference]{style/acmsiggraph}

\TOGonlineid{45678}
\TOGvolume{0}
\TOGnumber{0}
\TOGarticleDOI{1111111.2222222}
\TOGprojectURL{}
\TOGvideoURL{}
\TOGdataURL{}
\TOGcodeURL{}

\title{The Title of Your Paper Goes Here}

\author{Robert A. Smith\thanks{e-mail:rsmith@gmail.com}\\Smith Research}
\pdfauthor{Robert A. Smith}

\keywords{radiosity, global illumination, constant time}

\begin{document}

%% \teaser{
%%   \includegraphics[height=1.5in]{images/sampleteaser}
%%   \caption{Spring Training 2009, Peoria, AZ.}
%% }

\maketitle

\begin{abstract}

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
molestie consequat, vel illum dolore eu feugiat nulla facilisis at 
vero eros et accumsan et iusto odio dignissim qui blandit praesent
luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat.

Citations can be done this way~\cite{Jobs95} or this more concise 
way~\shortcite{Jobs95}, depending upon the application.

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit~\cite{Pellacini:2005:LAH}
in vulputate velit esse molestie~\cite{notes2002} 
consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
delenit augue duis dolore te feugait nulla facilisi.~\cite{Park:2006:DSI}

\end{abstract}

\begin{CRcatlist}
  \CRcat{I.3.3}{Computer Graphics}{Three-Dimensional Graphics and Realism}{Display Algorithms}
  \CRcat{I.3.7}{Computer Graphics}{Three-Dimensional Graphics and Realism}{Radiosity};
\end{CRcatlist}

\keywordlist

%% Use this only if you're preparing a technical paper to be published in the 
%% ACM 'Transactions on Graphics' journal.

\TOGlinkslist

%% Required for all content.

\copyrightspace

\section{Introduction}

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

\section{Exposition}

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

\begin{equation}
 \sum_{j=1}^{z} j = \frac{z(z+1)}{2}
\end{equation}

\begin{eqnarray}
x & \ll & y_{1} + \cdots + y_{n} \\
  & \leq & z
\end{eqnarray}

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

\section{Exposition}

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

\section*{Acknowledgements}

To Robert, for all the bagels.

\bibliographystyle{style/acmsiggraph}
\bibliography{paper}
\end{document}

### 2013-12-16 paper/template.tex
Update paper directory

%%% template.tex
%%%
%%% This LaTeX source document can be used as the basis for your technical
%%% paper or abstract. Intentionally stripped of annotation, the parameters
%%% and commands should be adjusted for your particular paper - title, 
%%% author, article DOI, etc.
%%% The accompanying ``template.annotated.tex'' provides copious annotation
%%% for the commands and parameters found in the source document. (The code
%%% is identical in ``template.tex'' and ``template.annotated.tex.'')

\documentclass[conference]{acmsiggraph}

\TOGonlineid{45678}
\TOGvolume{0}
\TOGnumber{0}
\TOGarticleDOI{1111111.2222222}
\TOGprojectURL{}
\TOGvideoURL{}
\TOGdataURL{}
\TOGcodeURL{}

\title{Interactive Visualization for System Log Analysis}

\author{Woojong Koh and Armin Samii\thanks{e-mail:\{wjkoh,samii\}@berkeley.edu}\\University of California, Berkeley}
\pdfauthor{Armin Samii and Woojong Koh}

\keywords{visualization, log, data mining, clustering}

\begin{document}

\teaser{
    \includegraphics[width=1.0\textwidth]{images/screenshot_full.png}
    \caption{Screenshot of Our Visualization}
}

\maketitle

\begin{abstract}

In this paper we present the way of visualizing large amount of system log data into single screen.
These log files from distributed systems can be tens of or hundreds of megabytes, and impossible to
parse without tools. Due to the amount of data we cluster machines by similar trends and also
pre-process the data in order to achieve real-time clustering according to user inputs. By doing so
we end up separating actual anomalous nodes from normal ones and can show the information in
visualization clearly.

\end{abstract}

\begin{CRcatlist}
  \CRcat{I.3.3}{Computer Graphics}{Three-Dimensional Graphics and Realism}{Display Algorithms}
  %CRcat{I.3.7}{Computer Graphics}{Three-Dimensional Graphics and Realism}{Radiosity};
\end{CRcatlist}

\keywordlist

%% Use this only if you're preparing a technical paper to be published in the 
%% ACM 'Transactions on Graphics' journal.

\TOGlinkslist

%% Required for all content.

\copyrightspace

\section{Introduction}
% An explanation of the problem and the motivation for solving it.

The gears of a distributed system are constantly moving, and the system administrator uses log files
to understand what goes on underneath. They are also spread out over many machines and many places in
these machines.

Our goal is to aggregate this data into a visualization that can be easily checked by the system
administrator. It should take no more than a glance to understand the system state, failures, or any
other desired information. This will allow the system administrator to regularly check the state
without setting up special scripts which parse the logs. For example, the system administrator
should be able to view the progress of a task while away from the office and without access to a
command-line.

The research problem is figuring out how to display megabytes of log data into a visually and
interactively simple interface and spot abnormalities without any additional effort.

With a large enough cluster, this will require focusing the user’s attention on “interesting” parts
of the system state. This will require detecting anomalies, which in turn requires a
constantly-evolving model of what the “normal” state of the system is.

\section{Related Works}
% A description of previous papers related to your project.

\begin{figure*}[p]
    \centering
    \includegraphics[width=0.9\textwidth]{images/ganglia.png}
    \caption{Screenshot of Ganglia\protect\footnotemark}
    \label{fig:ganglia}
\end{figure*}
\footnotetext{\url{http://en.wikipedia.org/wiki/File:ScalableGridEngineGanglia2.png}}
\begin{figure*}[p]
    \centering
    \includegraphics[width=0.9\textwidth]{images/rrdtool.png}
    \caption{Screenshot of RRDtool\protect\footnotemark}
    \label{fig:rrdtool}
\end{figure*}
\footnotetext{\url{http://oss.oetiker.ch/rrdtool/gallery/index.en.html}}

There are many good system monitor solutions out there including Ganglia~\cite{Massie04} and Munin
which is based on RRDtool~\cite{Oetiker99}. However, as you can see in Figure~\ref{fig:ganglia} and
Figure~\ref{fig:rrdtool}, the visualizations become harder to detect anomalies in the system as the
number of nodes increases.

We found Wei Xu and collaborators~\cite{Xu09} made very good system of detecting anomalies by mining
log data. We adopted their machine learning approach including their feature specifications in order
to create an intelligent visualization system.

\section{Methods}
% A detailed explanation of the techniques and algorithms you used to solve the problem.

We have a system which generates such log files using the Google Logging Library
(glog)\footnote{\url{https://code.google.com/p/google-glog/}}. Our goal is to create a generic
visualization library for glog, with emphasis on this system. Theoretically, anybody using glog in a
distributed framework should be able to work with the visualizations out-of-the-box. Here is the
summary of our algorithm:

\paragraph{Algorithm Summary}
\begin{enumerate}
    \item Compute per-node features
        \begin{itemize}
            \item Message count vector
            \item Statistics across time
            \item Each node computes its own features
        \end{itemize}
    \item Select k most representative nodes
        \begin{itemize}
            \item Using distributed k-means clustering
            \item User interactively chooses k
        \end{itemize}
    \item Display interactive visualization
        \begin{itemize}
            \item Client processes visualization
            \item Web server clusters on-demand
            \item Server communicates through Apache server
        \end{itemize}
\end{enumerate}

\subsection{Log Parser}
We mainly used Python for implementing our log parsing system. Google Logging Library provides a
consistent log format so we were able to build a parser easily using Python's regular expression
library. After parsing we need to group every log message by its type and extract nominal and
quantitative data from the messages.

First of all, we group log messages by its source file name and line number. After grouping them
into same types, we could separate dynamic parts of messages from static parts using text difference
algorithms. In Figure~\ref{fig:extracting_data}, you can see an example of extracting data. Since
the dynamic part of log messages will change over time, we can tell the exact part if we have a
large enough amount of log messages of the same type. After detecting dynamic parts, we can simply
parse the part of message strings into integer or floating number if quantitative and string if
nominal.

\subsection{Clusterer}
We also used Python, such as NumPy/SciPy and scikit-learn, for implementing our clustering system.
Due to the huge amount of data we need to preprocess the data in some way to boost the speed of the
clustering up to an interactive rate. As for clustering we used simple k-means clustering
\cite{kmeans,Lloyd82} of the scikit-learn library \cite{scikit-learn} and it turned out the
computation speed is quite good. However, the problem is in the feature extraction. We want to
extract features only from the user-specified time range, i.e., we want to apply clustering on only
features within the range. The range will change continuously by user inputs so we need to be able
to calculate features in real time in order to achieve our goal, an interactive visualization.

We mainly used a message count vector~\cite{Xu09} for our feature vectors and we change the domain
in most cases just by time. Thus, we first came up with data sorted in terms of time. However, just
sorted data structure couldn't help that much because the size of data is still large and it doesn't
fit into the main memory at all. Thus, the data should be stored in a secondary storage, which means
a hard disk drive. Due to the poor performance of random access of hard disk drives we couldn't get
much improvements using the sorted data structure.

Thus, to optimize further we introduced prefix sums of feature vectors. Prefix sums can help
calculating partial sums because we just need to subtract the first element from the last element of
the range to get a sum of values within the range.  We may use Fenwick tree \cite{Fenwick94} or
other advanced data structures but we satisfied with our current performance, which took about 0.4s
per each clustering.

\subsection{Visualization}
\begin{figure*}[p]
    \centering
    \includegraphics[width=0.9\textwidth]{images/screenshot_plot.png}
    \caption{Data points and raw log messages}
    \label{fig:ss_plot}
\end{figure*}

\begin{figure}[p]
    \centering
    \includegraphics[width=0.7\columnwidth]{images/screenshot_controls.png}
    \caption{User controls}
    \label{fig:ss_controls}
\end{figure}

\begin{figure}[p]
    \centering
    \includegraphics[width=1.0\columnwidth]{images/extracting_data_1.pdf}
    \includegraphics[width=1.0\columnwidth]{images/extracting_data_2.pdf}
    \caption{Text processing}
    \label{fig:extracting_data}
\end{figure}

We used D3~\cite{D311} and jQuery UI libraries for our user interfaces. The front end of our
framework is a web page so that you can explore it using any web browser supporting JavaScript and
SVG. Thus, we can say that we achieved one of our project goals, which is that implementing a system
that easily accessible from remote location using even mobile devices such as a smartphone.

In Figure \ref{fig:ss_controls} you can see the user controls of our visualization. The controls
basically consist of three sliders and one text box and each of them give an user a degree of
freedom in each dimension, such as time, the number of clusters, and the number of message types.

We used one range control for setting time range and two slider control for setting the number of
clusters to find and the number of message types to display, respectively. This system can response
to user inputs at an interactive rate because we speed up the clustering process and now it takes
only a half second as mentioned earlier.

\section{Results}
% The visualizations your system produces and data to help evaluate your approach. For example you
% may include running times, or the time users typically spend generating a visualization using your
% system.

\subsection{Anomaly Detection}
We set the metric for anomaly detection as follow: How many clusters are required until anomalous
node in its own cluster? In the best case of this scenario the number of cluster always should be
two if there is only one type of anomalies at a time.

In Figure~\ref{fig:anomaly} you can see the performance of our anomaly detection and it shows that
our system can tell anomalies such as a failing or slow node under the interference like random
noises. The system works well even when 50\% of the data is garbage, i.e., randomly generated logs.

\begin{figure}[p]
    \centering
    \includegraphics[width=1.0\columnwidth]{images/anomaly.pdf}
    \caption{Anomaly detection}
    \label{fig:anomaly}
\end{figure}

If we were able to find a ground-truth log file with labeled anomalies, we could compare our error
rates to those presented in papers by Wei Xu et al. Unfortunately, we couldn't find any publicly
accessible log database.

\subsection{Timing Information}
We ran on 500,000 messages per node and on average 30MB of data were sent to the HTTP Server per
node. Visualization took on the client side so the computation cost of visualization doesn't slow
down the server. In Figure~\ref{fig:timings} you can see basic timing information of our system.

\begin{figure}[p]
    \centering
    \includegraphics[width=1.0\columnwidth]{images/timing.pdf}
    \caption{Timing information}
    \label{fig:timings}
\end{figure}

\section{Discussion}
% What has the audience learned about visualization from your work?

The log data we used in this project came from the system which looks through terabytes of Google
Street View data and uses them for various machine learning tasks. It uses about 80 processes, each
process with its own log file, and spread over around a dozen machines. There are regular failures
which are currently cumbersome to find and debug.

We wanted to create a responsive system which visualizes the state of the system in real-time, is
robust to errors, and can display an arbitrary number of nodes. Not all nodes will be displayable on
a single screen, so the system had to be able to determine areas of interest (e.g. failed nodes,
slow nodes, or completed tasks) in the visualization and direct the user’s attention there. Our
anomaly detection built on Wei Xu's in several ways: (1) their work does not use timestamps, which
we believe are vital to understanding the system’s state; (2) their work intentionally ignores the
ordering of log messages, and instead pays attention only to the number of occurrences of a message
, we hypothesize order may matter; (3) most importantly, their visualizations are more primitive. We
believe the idea of near-infinite zoom (where the most detailed layer essentially shows the
information in the actual log, and as you zoom out, more and more information is aggregated) will
allow a user to explore the system state rapidly.

What we want to point out in this paper is that along with our project there is a direction of
visualization research that can ultimately help system administrators to monitor and analyze the
status of a large distributed system. We think this will become more and more important as the trend
of technology emphasizes on parallelism and distributed system such as cloud computing.

\section{Future Work}
% A description of how your system could be extended.}

We believe automatically determining plot type would be a good improvement for our system because
data extracted from raw log messages can be arbitrary and it is hard to be predetermined by an user.
Also, we want to extend our system to online and real-time analysis and implement more intelligent
features preprocessing. Lastly, it would be great if our system can cluster logs based on more
elaborate features.

\section*{Acknowledgements}
The authors appreciate Ling Huang for his helpful advices and Sean Arietta for providing the log
data gathered from his research project.

\bibliographystyle{acmsiggraph}
\bibliography{template}
\end{document}

### 2013-12-17 paper/paper.tex
Teaser updates

@@
 
-\title{The Title of Your Paper Goes Here}
+\title{Interactive Visualization for System Log Analysis}
 
-\author{Robert A. Smith\thanks{e-mail:rsmith@gmail.com}\\Smith Research}
-\pdfauthor{Robert A. Smith}
+\author{Woojong Koh and Armin Samii\thanks{e-mail:\{wjkoh,samii\}@berkeley.edu}\\University of California, Berkeley}
+\pdfauthor{Armin Samii and Woojong Koh}
 
-\keywords{radiosity, global illumination, constant time}
+\keywords{visualization, log, data mining, clustering}
 
@@
 
-%% \teaser{
-%%   \includegraphics[height=1.5in]{images/sampleteaser}
-%%   \caption{Spring Training 2009, Peoria, AZ.}
-%% }
+\teaser{
+    \includegraphics[width=1.0\textwidth]{images/screenshot_full.png}
+    \caption{Our visualization displays arbitrary system log data by parsing the variable parts of each log message.
+Each row contains messages from a single node.
+Each column is a certain type of message.
+Each point is a single log message.
+The x axis is time; the y axis is the value of the extracted variable.
+The colors indicate multiple variables in the same message.
+For example, column two has two colors: blue is the number of valid patches, and red is the number of invalid patches.
+Notice here that five nodes performed similarly (row 2), whereas one node (row 1) is performing more slowly, and thus the number of log messages is sparse.
+}
+}
 
@@
 
-Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
-molestie consequat, vel illum dolore eu feugiat nulla facilisis at 
-vero eros et accumsan et iusto odio dignissim qui blandit praesent
-luptatum zzril delenit augue duis dolore te feugait nulla
-facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
-sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
-aliquam erat volutpat.
-
-Citations can be done this way~\cite{Jobs95} or this more concise 
-way~\shortcite{Jobs95}, depending upon the application.
-
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
-vel eum iriure dolor in hendrerit~\cite{Pellacini:2005:LAH}
-in vulputate velit esse molestie~\cite{notes2002} 
-consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
-accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
-delenit augue duis dolore te feugait nulla facilisi.~\cite{Park:2006:DSI}
+In this paper we present the way of visualizing large amount of system log data into single screen.
+These log files from distributed systems can be tens of or hundreds of megabytes, and impossible to
+parse without tools. Due to the amount of data we cluster machines by similar trends and also
+pre-process the data in order to achieve real-time clustering according to user inputs. By doing so
+we end up separating actual anomalous nodes from normal ones and can show the information in
+visualization clearly.
 
@@
   \CRcat{I.3.3}{Computer Graphics}{Three-Dimensional Graphics and Realism}{Display Algorithms}
-  \CRcat{I.3.7}{Computer Graphics}{Three-Dimensional Graphics and Realism}{Radiosity};
+  %CRcat{I.3.7}{Computer Graphics}{Three-Dimensional Graphics and Realism}{Radiosity};
 \end{CRcatlist}
@@
 \section{Introduction}
+% An explanation of the problem and the motivation for solving it.
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
+The gears of a distributed system are constantly moving, and the system administrator uses log files
+to understand what goes on underneath. They are also spread out over many machines and many places in
+these machines.
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
+Our goal is to aggregate this data into a visualization that can be easily checked by the system
+administrator. It should take no more than a glance to understand the system state, failures, or any
+other desired information. This will allow the system administrator to regularly check the state
+without setting up special scripts which parse the logs. For example, the system administrator
+should be able to view the progress of a task while away from the office and without access to a
+command-line.
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
+The research problem is figuring out how to display megabytes of log data into a visually and
+interactively simple interface and spot abnormalities without any additional effort.
 
-\section{Exposition}
+With a large enough cluster, this will require focusing the user’s attention on “interesting” parts
+of the system state. This will require detecting anomalies, which in turn requires a
+constantly-evolving model of what the “normal” state of the system is.
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
+\section{Related Works}
+% A description of previous papers related to your project.
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
+\begin{figure*}[p]
+    \centering
+    \includegraphics[width=0.9\textwidth]{images/ganglia.png}
+    \caption{Screenshot of Ganglia\protect\footnotemark}
+    \label{fig:ganglia}
+\end{figure*}
+\footnotetext{\url{http://en.wikipedia.org/wiki/File:ScalableGridEngineGanglia2.png}}
+\begin{figure*}[p]
+    \centering
+    \includegraphics[width=0.9\textwidth]{images/rrdtool.png}
+    \caption{Screenshot of RRDtool\protect\footnotemark}
+    \label{fig:rrdtool}
+\end{figure*}
+\footnotetext{\url{http://oss.oetiker.ch/rrdtool/gallery/index.en.html}}
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
+There are many good system monitor solutions out there including Ganglia~\cite{Massie04} and Munin
+which is based on RRDtool~\cite{Oetiker99}. However, as you can see in Figure~\ref{fig:ganglia} and
+Figure~\ref{fig:rrdtool}, the visualizations become harder to detect anomalies in the system as the
+number of nodes increases.
 
-\begin{equation}
- \sum_{j=1}^{z} j = \frac{z(z+1)}{2}
-\end{equation}
+We found Wei Xu and collaborators~\cite{Xu09} made very good system of detecting anomalies by mining
+log data. We adopted their machine learning approach including their feature specifications in order
+to create an intelligent visualization system.
 
-\begin{eqnarray}
-x & \ll & y_{1} + \cdots + y_{n} \\
-  & \leq & z
-\end{eqnarray}
+\section{Methods}
+% A detailed explanation of the techniques and algorithms you used to solve the problem.
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
+We have a system which generates such log files using the Google Logging Library
+(glog)\footnote{\url{https://code.google.com/p/google-glog/}}. Our goal is to create a generic
+visualization library for glog, with emphasis on this system. Theoretically, anybody using glog in a
+distributed framework should be able to work with the visualizations out-of-the-box. Here is the
+summary of our algorithm:
 
-\section{Exposition}
+\paragraph{Algorithm Summary}
+\begin{enumerate}
+    \item Compute per-node features
+        \begin{itemize}
+            \item Message count vector
+            \item Statistics across time
+            \item Each node computes its own features
+        \end{itemize}
+    \item Select k most representative nodes
+        \begin{itemize}
+            \item Using distributed k-means clustering
+            \item User interactively chooses k
+        \end{itemize}
+    \item Display interactive visualization
+        \begin{itemize}
+            \item Client processes visualization
+            \item Web server clusters on-demand
+            \item Server communicates through Apache server
+        \end{itemize}
+\end{enumerate}
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
+\subsection{Log Parser}
+We mainly used Python for implementing our log parsing system. Google Logging Library provides a
+consistent log format so we were able to build a parser easily using Python's regular expression
+library. After parsing we need to group every log message by its type and extract nominal and
+quantitative data from the messages.
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
+First of all, we group log messages by its source file name and line number. After grouping them
+into same types, we could separate dynamic parts of messages from static parts using text difference
+algorithms. In Figure~\ref{fig:extracting_data}, you can see an example of extracting data. Since
+the dynamic part of log messages will change over time, we can tell the exact part if we have a
+large enough amount of log messages of the same type. After detecting dynamic parts, we can simply
+parse the part of message strings into integer or floating number if quantitative and string if
+nominal.
+
+\subsection{Clusterer}
+We also used Python, such as NumPy/SciPy and scikit-learn, for implementing our clustering system.
+Due to the huge amount of data we need to preprocess the data in some way to boost the speed of the
+clustering up to an interactive rate. As for clustering we used simple k-means clustering
+\cite{kmeans,Lloyd82} of the scikit-learn library \cite{scikit-learn} and it turned out the
+computation speed is quite good. However, the problem is in the feature extraction. We want to
+extract features only from the user-specified time range, i.e., we want to apply clustering on only
+features within the range. The range will change continuously by user inputs so we need to be able
+to calculate features in real time in order to achieve our goal, an interactive visualization.
+
+We mainly used a message count vector~\cite{Xu09} for our feature vectors and we change the domain
+in most cases just by time. Thus, we first came up with data sorted in terms of time. However, just
+sorted data structure couldn't help that much because the size of data is still large and it doesn't
+fit into the main memory at all. Thus, the data should be stored in a secondary storage, which means
+a hard disk drive. Due to the poor performance of random access of hard disk drives we couldn't get
+much improvements using the sorted data structure.
+
+Thus, to optimize further we introduced prefix sums of feature vectors. Prefix sums can help
+calculating partial sums because we just need to subtract the first element from the last element of
+the range to get a sum of values within the range.  We may use Fenwick tree \cite{Fenwick94} or
+other advanced data structures but we satisfied with our current performance, which took about 0.4s
+per each clustering.
+
+\subsection{Visualization}
+\begin{figure*}[p]
+    \centering
+    \includegraphics[width=0.9\textwidth]{images/screenshot_plot.png}
+    \caption{Data points and raw log messages}
+    \label{fig:ss_plot}
+\end{figure*}
+
+\begin{figure}[p]
+    \centering
+    \includegraphics[width=0.7\columnwidth]{images/screenshot_controls.png}
+    \caption{User controls}
+    \label{fig:ss_controls}
+\end{figure}
+
+\begin{figure}[p]
+    \centering
+    \includegraphics[width=1.0\columnwidth]{images/extracting_data_1.pdf}
+    \includegraphics[width=1.0\columnwidth]{images/extracting_data_2.pdf}
+    \caption{Text processing}
+    \label{fig:extracting_data}
+\end{figure}
+
+We used D3~\cite{D311} and jQuery UI libraries for our user interfaces. The front end of our
+framework is a web page so that you can explore it using any web browser supporting JavaScript and
+SVG. Thus, we can say that we achieved one of our project goals, which is that implementing a system
+that easily accessible from remote location using even mobile devices such as a smartphone.
+
+In Figure \ref{fig:ss_controls} you can see the user controls of our visualization. The controls
+basically consist of three sliders and one text box and each of them give an user a degree of
+freedom in each dimension, such as time, the number of clusters, and the number of message types.
+
+We used one range control for setting time range and two slider control for setting the number of
+clusters to find and the number of message types to display, respectively. This system can response
+to user inputs at an interactive rate because we speed up the clustering process and now it takes
+only a half second as mentioned earlier.
+
+\section{Results}
+% The visualizations your system produces and data to help evaluate your approach. For example you
+% may include running times, or the time users typically spend generating a visualization using your
+% system.
+
+\subsection{Anomaly Detection}
+We set the metric for anomaly detection as follow: How many clusters are required until anomalous
+node in its own cluster? In the best case of this scenario the number of cluster always should be
+two if there is only one type of anomalies at a time.
+
+In Figure~\ref{fig:anomaly} you can see the performance of our anomaly detection and it shows that
+our system can tell anomalies such as a failing or slow node under the interference like random
+noises. The system works well even when 50\% of the data is garbage, i.e., randomly generated logs.
+
+\begin{figure}[p]
+    \centering
+    \includegraphics[width=1.0\columnwidth]{images/anomaly.pdf}
+    \caption{Anomaly detection}
+    \label{fig:anomaly}
+\end{figure}
+
+If we were able to find a ground-truth log file with labeled anomalies, we could compare our error
+rates to those presented in papers by Wei Xu et al. Unfortunately, we couldn't find any publicly
+accessible log database.
+
+\subsection{Timing Information}
+We ran on 500,000 messages per node and on average 30MB of data were sent to the HTTP Server per
+node. Visualization took on the client side so the computation cost of visualization doesn't slow
+down the server. In Figure~\ref{fig:timings} you can see basic timing information of our system.
+
+\begin{figure}[p]
+    \centering
+    \includegraphics[width=1.0\columnwidth]{images/timing.pdf}
+    \caption{Timing information}
+    \label{fig:timings}
+\end{figure}
+
+\section{Discussion}
+% What has the audience learned about visualization from your work?
+
+The log data we used in this project came from the system which looks through terabytes of Google
+Street View data and uses them for various machine learning tasks. It uses about 80 processes, each
+process with its own log file, and spread over around a dozen machines. There are regular failures
+which are currently cumbersome to find and debug.
+
+We wanted to create a responsive system which visualizes the state of the system in real-time, is
+robust to errors, and can display an arbitrary number of nodes. Not all nodes will be displayable on
+a single screen, so the system had to be able to determine areas of interest (e.g. failed nodes,
+slow nodes, or completed tasks) in the visualization and direct the user’s attention there. Our
+anomaly detection built on Wei Xu's in several ways: (1) their work does not use timestamps, which
+we believe are vital to understanding the system’s state; (2) their work intentionally ignores the
+ordering of log messages, and instead pays attention only to the number of occurrences of a message
+, we hypothesize order may matter; (3) most importantly, their visualizations are more primitive. We
+believe the idea of near-infinite zoom (where the most detailed layer essentially shows the
+information in the actual log, and as you zoom out, more and more information is aggregated) will
+allow a user to explore the system state rapidly.
+
+What we want to point out in this paper is that along with our project there is a direction of
+visualization research that can ultimately help system administrators to monitor and analyze the
+status of a large distributed system. We think this will become more and more important as the trend
+of technology emphasizes on parallelism and distributed system such as cloud computing.
+
+\section{Future Work}
+% A description of how your system could be extended.}
+
+We believe automatically determining plot type would be a good improvement for our system because
+data extracted from raw log messages can be arbitrary and it is hard to be predetermined by an user.
+Also, we want to extend our system to online and real-time analysis and implement more intelligent
+features preprocessing. Lastly, it would be great if our system can cluster logs based on more
+elaborate features.
 
 \section*{Acknowledgements}
-
-To Robert, for all the bagels.
+The authors appreciate Ling Huang for his helpful advices and Sean Arietta for providing the log
+data gathered from his research project.
 

### 2013-12-17 paper/paper.tex
Updated intro and abstract

@@
 
-In this paper we present the way of visualizing large amount of system log data into single screen.
-These log files from distributed systems can be tens of or hundreds of megabytes, and impossible to
-parse without tools. Due to the amount of data we cluster machines by similar trends and also
-pre-process the data in order to achieve real-time clustering according to user inputs. By doing so
-we end up separating actual anomalous nodes from normal ones and can show the information in
-visualization clearly.
+We present a visualization interface to assist system administrators in searching system logs.
+Even with current regular expression matchers such as \texttt{grep}, the amount of log data is difficult for humans to understand.
+We augment \texttt{grep} with a visualization of the matched patterns.
+To further reduce the amount of information displayed, we cluster similarly-performing nodes and only show a single node representative of the entire cluster.
+The user can then interactively search the log, zoom in on a certain time window, and choose the number of clusters.
+%This means that at the most ``zoomed out'' state (no time window specified, no search patters, and no clustering), the raw log data is displayed.
+Our simulations show that our clustering is effective and our system is fast:
+the clustering successfully isolates anomalously-performing nodes in a variety of situations;
+the visualization can interactively visualize hundreds of millions of log messages across a hundred nodes in real-time.
+TODO: Get the numbers from this simulation.
 
@@
 \section{Introduction}
-% An explanation of the problem and the motivation for solving it.
 
-The gears of a distributed system are constantly moving, and the system administrator uses log files
-to understand what goes on underneath. They are also spread out over many machines and many places in
-these machines.
+The gears of a distributed system are complex and constantly moving.
+A system administrator uses log files to understand the underlying structure and detect anomalous behavior.
+These logs are large, distributed across many filesystems, and hard to understand even with today's tools.
 
-Our goal is to aggregate this data into a visualization that can be easily checked by the system
-administrator. It should take no more than a glance to understand the system state, failures, or any
-other desired information. This will allow the system administrator to regularly check the state
-without setting up special scripts which parse the logs. For example, the system administrator
-should be able to view the progress of a task while away from the office and without access to a
-command-line.
-
-The research problem is figuring out how to display megabytes of log data into a visually and
-interactively simple interface and spot abnormalities without any additional effort.
+Our goal is to aggregate the log data into a visualization that can be easily understood by the system administrator.
+The two primary use cases are:
+\begin{enumerate}
+\item searching through system logs to find software bugs or hardware failures, and
+\item viewing the current system state on-the-go without access to a command line.
+\end{enumerate}
+To this end, our system contributes the following features:
+\begin{enumerate}
+\item \textbf{trustable}: the user is able to view the raw data at the most ``zoomed out'' state;
+\item \textbf{fast and distributed}: the proprocessing primarily occurs at each node, with a single central node aggregating preprocessed data; and
+\item \textbf{interactive}: the user can adjust the paramaters of the visualization at interactive speeds.
+\end{enumerate}
 
@@
 
-\section{Related Works}
+\section{Related Work}
 % A description of previous papers related to your project.

### 2013-12-17 paper/paper.tex
Updated related work, began algorithm section

@@
 
+Our system is open source and available on github~\footnote{\url{http://www.github.com/wjkoh/cs262a}}.
+
 \section{Related Work}
-% A description of previous papers related to your project.
+Given the wide range of distributed applications, system log analysis has been a prominent research area for decades.
+Unfortunately, given the available visualization tools in the past, none are adequate for off-the-shelf use.
+We divide the related work into two sections: system state analysis and system log analysis.
+The system state refers to aggregating data about the statistics of resource usage in a system, agnostic to the application.
+System logs are arbitrary print statements in a piece of code, distributed across one or more nodes.
 
-\begin{figure*}[p]
+\subsection{System State Analysis}
+Ganglia\footnote{\url{http://www.ganglia.info}} is a system monitoring system which can handle up to 2,000 nodes and aggregate statistics in real-time~\cite{Massie04}.
+It requires installing software on each running node and can only handle predefined metrics; it does not allow user-defined metrics and does not utilize system logs.
+The data is not clustered either, so either all nodes are visualized, or all nodes are aggregated.
+Finally, it is far from interactive: simple options such as setting a time window require many clicks.
+See Figure~\ref{fig:ganglia} for a screenshot of the amount of data shown for a single node.
+
+\begin{figure*}[h]
     \centering
     \includegraphics[width=0.9\textwidth]{images/ganglia.png}
-    \caption{Screenshot of Ganglia\protect\footnotemark}
+    \caption{Screenshot of Ganglia (image courtesy of Wikipedia). This information shows various metrics for a single node.}
     \label{fig:ganglia}
 \end{figure*}
-\footnotetext{\url{http://en.wikipedia.org/wiki/File:ScalableGridEngineGanglia2.png}}
+
 \begin{figure*}[p]
@@
 
-There are many good system monitor solutions out there including Ganglia~\cite{Massie04} and Munin
-which is based on RRDtool~\cite{Oetiker99}. However, as you can see in Figure~\ref{fig:ganglia} and
-Figure~\ref{fig:rrdtool}, the visualizations become harder to detect anomalies in the system as the
-number of nodes increases.
+Munin, based on RRDtool, is similar to Ganglia: after installation on each node, it automatically presents a visualization of the system status~\cite{Oetiker99}.
+As is evident from Figure~\ref{fig:rrdtool}, the visualizations is extremely dense and illegible as the number of nodes increase.
 
-We found Wei Xu and collaborators~\cite{Xu09} made very good system of detecting anomalies by mining
-log data. We adopted their machine learning approach including their feature specifications in order
-to create an intelligent visualization system.
+We use ideas from these system state visualizations, and apply them to system logs.
+We also recognize that viewing all logs makes an illegible visualization, viewing one node at a time is inefficient, and aggregating all nodes into a single server status hides any anomalous behavior.
+
+\subsection{System Log Analysis}
+
+Wei Xu and collaborators~\cite{Xu09} uses machine learning to detect anomalies by mining log data.
+When an anomaly is detected, a decision tree visualization is presented to the user which attempts to explain why it is an anomaly.
+This setup, while useful, is fundamentally untrustable: no algorithm can successfully detect all anomalies, and the user will eventually have to resort to command-line \texttt{grep}.
+We adopt their machine learning features for our clustering algorithm.
+
+HP Operations Analytics\footnote{\url{http://www8.hp.com/us/en/software-solutions/operations-analytics-operations-analysis/}} provides many similar features to our work, including interactive time-window zooming.
+It does not seem to cluster any data, nor does it focus on software bugs, but instead analyze single-node server issues.
+It seems to require modifying source code to work with their log analysis system, which may limit its applicabilty.
+Unfortunately, it is also closed-source proprietary software so we are not able to test out its effectiveness.
+
+Seaview has the closest goals to ours~\cite{hangalseaview}, and provide similar visualizations.
+However, they do not work with arbitrary log types (notably, log messages with multiple variables), nor does it extend to multiple nodes or large amounts of data.
+They include no machine learning or reduction of the amount of data shown, other than a planned \texttt{grep} text search.
+
+Makanju and collaborators present spatio-temporal clustering of log data as well \cite{makanju2012spatio}.
+Their method does not seem to be, nor claim to be, interactive.
+We find their clustering method and features to be complementary to ours, and it would fit cleanly in our interactive system.
 
 \section{Methods}
-% A detailed explanation of the techniques and algorithms you used to solve the problem.
 
-We have a system which generates such log files using the Google Logging Library
-(glog)\footnote{\url{https://code.google.com/p/google-glog/}}. Our goal is to create a generic
-visualization library for glog, with emphasis on this system. Theoretically, anybody using glog in a
-distributed framework should be able to work with the visualizations out-of-the-box. Here is the
-summary of our algorithm:
+We automatically extract the source code line which printed each log message, the time it was printed, the node which generated it, and the variable part of each log message.
 
-\paragraph{Algorithm Summary}
+Our algorithm can be summarized as follows:
+
 \begin{enumerate}
     \item Compute per-node features
-        \begin{itemize}
-            \item Message count vector
-            \item Statistics across time
-            \item Each node computes its own features
-        \end{itemize}
-    \item Select k most representative nodes
-        \begin{itemize}
-            \item Using distributed k-means clustering
-            \item User interactively chooses k
-        \end{itemize}
+    \item Perform distributed k-means clustering
     \item Display interactive visualization
-        \begin{itemize}
-            \item Client processes visualization
-            \item Web server clusters on-demand
-            \item Server communicates through Apache server
-        \end{itemize}
 \end{enumerate}
+
+We desribe each of these in detail below.
+
+\subsection{Feature Extraction}
+Each node computes its own features, and passes on the feature vector as well as raw log data to the master server.
+
+Following the example of Xu and collaborators \cite{Xu09}, we extract a message count vector in the time window requested by the user.
+The message count vector is the number of unique times each message was printed.
+This is a strong feature for failing or slowing nodes.
+The state ratio vector compares how many times one type of message was printed compared to similar ones; for example, the ratio of file opens to file closes.
+We found that the state ratio vector was not useful for our log data, and so we do not compute it to save CPU time.
+
+Using the extracted variables, we compute basic statistics about each type of message.
+In a given time window, we compute the:
+\begin{enumerate}
+\item mean,
+\item standard deviation,
+\item slope of a linear regression,
+\item z-score, and
+\item r-value.
+\end{enumerate}
+These would indicate anomalies involving, for example, a diverging residual in a machine learning application, a section of code which takes too long to run, and a message which keeps working on the same file rather than working on new files.
+
+\subsection{Clustering}
+    %\item Perform distributed k-means clustering
+
+\subsection{Visualization}
+    %\item Display interactive visualization
+
+%         \begin{itemize}
+%             \item Message count vector
+%             \item Statistics across time
+%             \item Each node computes its own features
+%         \end{itemize}
+%     \item Select k most representative nodes
+%         \begin{itemize}
+%             \item Using distributed k-means clustering
+%             \item User interactively chooses k
+%         \end{itemize}
+%     \item Display interactive visualization
+%         \begin{itemize}
+%             \item Client processes visualization
+%             \item Web server clusters on-demand
+%             \item Server communicates through Apache server
+%         \end{itemize}
+%
+
+\section{Implementation}
+We identify each log message by the node and source code line that generated it.
+We assume that all logs are generated using the Google Logging Library (GLOG)\footnote{\url{https://code.google.com/p/google-glog/}} in order to know which source line generated which log message.
+This assumption can be dropped using the mining algorithm of Xu and collaborators~\cite{Xu:2009:OSP:1674659.1677125}.
 

### 2013-12-17 paper/paper.tex
Updated implementation section

@@
 }
+    \label{fig:teaser}
 }
@@
 
-\begin{figure*}[h]
+\begin{figure*}[p]
     \centering
@@
     \includegraphics[width=0.9\textwidth]{images/rrdtool.png}
-    \caption{Screenshot of RRDtool\protect\footnotemark}
+    \caption{A screenshot of RRDtool attempting to coherently display a lot of information to the user, but ending up extremely cluttered.\protect\footnotemark}
     \label{fig:rrdtool}
@@
 \subsection{Clustering}
-    %\item Perform distributed k-means clustering
+We perform k-means clustering in the feature vector space~\cite{kmeans,Lloyd82}.
+With a small number of nodes (on the order of tens), the clustering can be performed on a single machine quickly.
+On more nodes, a distributed $k$-means algorithm is required, which can either be exact~\cite{jin2006fast} or approximate and slower~\cite{kanungo2002local}.
+Since the clustering will occur each time the user updates the parameters of the visualization, it is essential that it is fast.
+
+Once the desired number of centroids are found, we compute the node which is closest to the centroid in a least-squares sense and display only that node.
+Therefore, thhe visualization master server need only receive $O(k)$ data to pass on to the user, linear with the number of centroids, instead of with the total number of nodes.
 
 \subsection{Visualization}
-    %\item Display interactive visualization
+The master server collects the data from each node and uses an HTTP Apache Server to pass the data on to the client.
+We use d3.js to visualize the data~\cite{D311}.
+We use a scatterplot where each point is a raw log message, the x axis is time, and the y axis is the parsed variables (see Figure~\ref{fig:teaser}).
+Mousing over a scatter point displays the text of the raw log message.
+Each row in the visualization is the nearest node to the centroid of the $k$-means clustering.
+Each column is a specific type of log message, identified by its line in the source code.
 
-%         \begin{itemize}
-%             \item Message count vector
-%             \item Statistics across time
-%             \item Each node computes its own features
-%         \end{itemize}
-%     \item Select k most representative nodes
-%         \begin{itemize}
-%             \item Using distributed k-means clustering
-%             \item User interactively chooses k
-%         \end{itemize}
-%     \item Display interactive visualization
-%         \begin{itemize}
-%             \item Client processes visualization
-%             \item Web server clusters on-demand
-%             \item Server communicates through Apache server
-%         \end{itemize}
-%
+The user can interactively change the time window, the number of desired clusters, and the regular expression pattern.
+With each change, the clusters are recomputed and new data is shown.
+Clustering is only performed on the log messages which match the regular expression pattern in the given time window.
+
+Because the visualization gets cluttered if there are too many scatter points in a single graph, each graph at most shows 5,000 messages, each one evenly spaced in time.
+Zooming in (narrowing the time range) allows finer sampling of the log messages.
 
@@
 \subsection{Log Parser}
-We mainly used Python for implementing our log parsing system. Google Logging Library provides a
-consistent log format so we were able to build a parser easily using Python's regular expression
-library. After parsing we need to group every log message by its type and extract nominal and
+GLOG provides a consistent log format so we were able to build a parser easily using Python's regular expression
+library. After parsing we need to group every log message by its message type and extract nominal and
 quantitative data from the messages.
 
-First of all, we group log messages by its source file name and line number. After grouping them
-into same types, we could separate dynamic parts of messages from static parts using text difference
-algorithms. In Figure~\ref{fig:extracting_data}, you can see an example of extracting data. Since
-the dynamic part of log messages will change over time, we can tell the exact part if we have a
-large enough amount of log messages of the same type. After detecting dynamic parts, we can simply
-parse the part of message strings into integer or floating number if quantitative and string if
-nominal.
+First, we group log messages by its source file name and line number.
+We then separate dynamic parts of messages from static parts using Ratcliff/Obershelp pattern recognition \cite{pattermatch}.
+If the dynamic part is a string instead of number (such as a filename), we simply discretize the value.
+This will succeed in detecting if an anomalous node is opening the same file repeatedly, and other similar issues, but will fail on text data that is semantically meaningful.
 
-\subsection{Clusterer}
-We also used Python, such as NumPy/SciPy and scikit-learn, for implementing our clustering system.
-Due to the huge amount of data we need to preprocess the data in some way to boost the speed of the
-clustering up to an interactive rate. As for clustering we used simple k-means clustering
-\cite{kmeans,Lloyd82} of the scikit-learn library \cite{scikit-learn} and it turned out the
-computation speed is quite good. However, the problem is in the feature extraction. We want to
-extract features only from the user-specified time range, i.e., we want to apply clustering on only
-features within the range. The range will change continuously by user inputs so we need to be able
-to calculate features in real time in order to achieve our goal, an interactive visualization.
-
-We mainly used a message count vector~\cite{Xu09} for our feature vectors and we change the domain
-in most cases just by time. Thus, we first came up with data sorted in terms of time. However, just
-sorted data structure couldn't help that much because the size of data is still large and it doesn't
-fit into the main memory at all. Thus, the data should be stored in a secondary storage, which means
-a hard disk drive. Due to the poor performance of random access of hard disk drives we couldn't get
-much improvements using the sorted data structure.
-
-Thus, to optimize further we introduced prefix sums of feature vectors. Prefix sums can help
-calculating partial sums because we just need to subtract the first element from the last element of
-the range to get a sum of values within the range.  We may use Fenwick tree \cite{Fenwick94} or
-other advanced data structures but we satisfied with our current performance, which took about 0.4s
-per each clustering.
+\subsection{Feature Extraction and Clustering}
+The parameters requested by a user affect the input to the clustering, so we need to be able to compute feature extraction at interactive rates.
+Our current implementation only uses the message count vector~\cite{Xu09} for our feature vector to achieve such rates.
+Our message count vector computation uses prefix sums, where each message at time $t=T$ knows the message count in the range $t \in [0, T]$.
+To find the message count vector in a certain timerange $t \in [t0, t1]$, we find the closest message after $t=t0$ and the closest message before $t=t1$, and subtract the prefix sum at each of those times.
+This results in $t \in [0,t1] - [0,t0] = t \in [t0,t1]$, the exact range we want.
+Similar optimizations can be performed for the simple statistical features.
 
 \subsection{Visualization}
-\begin{figure*}[p]
-    \centering
-    \includegraphics[width=0.9\textwidth]{images/screenshot_plot.png}
-    \caption{Data points and raw log messages}
-    \label{fig:ss_plot}
-\end{figure*}
 
-\begin{figure}[p]
+We used d3.js~\cite{D311} and jQuery UI libraries for our user interfaces.
+The front end of our framework is a web-based dashboard, enabling easy viewing from a mobile device while on-the-go.
+However, in our experience, the interaction is much simpler on a large display.
+
+\begin{figure}[t]
     \centering
@@
 
-\begin{figure}[p]
-    \centering
-    \includegraphics[width=1.0\columnwidth]{images/extracting_data_1.pdf}
-    \includegraphics[width=1.0\columnwidth]{images/extracting_data_2.pdf}
-    \caption{Text processing}
-    \label{fig:extracting_data}
-\end{figure}
-
-We used D3~\cite{D311} and jQuery UI libraries for our user interfaces. The front end of our
-framework is a web page so that you can explore it using any web browser supporting JavaScript and
-SVG. Thus, we can say that we achieved one of our project goals, which is that implementing a system
-that easily accessible from remote location using even mobile devices such as a smartphone.
-
-In Figure \ref{fig:ss_controls} you can see the user controls of our visualization. The controls
-basically consist of three sliders and one text box and each of them give an user a degree of
-freedom in each dimension, such as time, the number of clusters, and the number of message types.
-
-We used one range control for setting time range and two slider control for setting the number of
-clusters to find and the number of message types to display, respectively. This system can response
-to user inputs at an interactive rate because we speed up the clustering process and now it takes
-only a half second as mentioned earlier.
+In Figure \ref{fig:ss_controls} you can see the user controls of our visualization.
+The controls allow control of the time window, the number of clusters, and the regular expression command.
+There is also control over the number of message types displayed, which should only display the most representative message types;
+we have not yet implemented this algorithm.
 
 \section{Results}
-% The visualizations your system produces and data to help evaluate your approach. For example you
-% may include running times, or the time users typically spend generating a visualization using your
-% system.
 

### 2013-12-17 paper/paper.tex
Updated results and style and format

@@
 
-\begin{CRcatlist}
-  \CRcat{I.3.3}{Computer Graphics}{Three-Dimensional Graphics and Realism}{Display Algorithms}
-  %CRcat{I.3.7}{Computer Graphics}{Three-Dimensional Graphics and Realism}{Radiosity};
-\end{CRcatlist}
-
-\keywordlist
-
 %% Use this only if you're preparing a technical paper to be published in the 
@@
 \TOGlinkslist
-
-%% Required for all content.
-
-\copyrightspace
 
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{images/ganglia.png}
+    \includegraphics[width=0.85\textwidth]{images/ganglia.png}
     \caption{Screenshot of Ganglia (image courtesy of Wikipedia). This information shows various metrics for a single node.}
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{images/rrdtool.png}
+    \includegraphics[width=0.85\textwidth]{images/rrdtool.png}
     \caption{A screenshot of RRDtool attempting to coherently display a lot of information to the user, but ending up extremely cluttered.\protect\footnotemark}
@@
     \includegraphics[width=0.7\columnwidth]{images/screenshot_controls.png}
-    \caption{User controls}
+    \caption{The configuration options available to the user via our web framework. The user can change these options at interactive rates.}
     \label{fig:ss_controls}
@@
 
+The success of our system depends on three axes: clustering performance, speed, and user friendliness.
+The third axis is the most difficult to evaluate, but we currently have one user of the system who is excited to use and improve it.
+We formally evaluate the other two axes here.
+
 \subsection{Anomaly Detection}
-We set the metric for anomaly detection as follow: How many clusters are required until anomalous
-node in its own cluster? In the best case of this scenario the number of cluster always should be
-two if there is only one type of anomalies at a time.
+\label{subsec:results:anomaly}
+The metric of success for anomaly detection is as follow: assuming only one node is anomalous, how many clusters are required until that node is in its own cluster?
+The best-case scenario is two clusters: one for each of the $n-1$ well-performing nodes, and another for the anomalous node.
 
-In Figure~\ref{fig:anomaly} you can see the performance of our anomaly detection and it shows that
-our system can tell anomalies such as a failing or slow node under the interference like random
-noises. The system works well even when 50\% of the data is garbage, i.e., randomly generated logs.
+We simulate five experiments on a seven node cluster using six nodes of real data and one node of simulated anomalous data.
+The five anomalies tested are:
+\begin{enumerate}
+\item
+\textbf{Failed node}
 
-\begin{figure}[p]
+One node fails when the algorithm is 80\% to completion, and all other nodes run to completion.
+\item
+\textbf{Node 5\% slower}
+
+One node runs 5\% slower than the speed of the other nodes
+\item
+\textbf{Node 1\% slower}
+
+One node runs 1\% slower than the speed of the other nodes
+\item
+\textbf{50\% of log is junk}
+
+Half of the messages in each log file are random, both in time and in value. One node is still 5\% slower. Can the anomalous node be detected?
+\item
+\textbf{99\% of log is junk}
+
+Only one out of every hundred messages is real. One node is still 5\% slower. Can the anomalous node be detected? Note that the filesize here is bloated to 100 times the actual log data.
+\end{enumerate}
+
+Figure~\ref{fig:anomaly} shows the performance of our clustering on each of the above tests.
+Most of the clustering successfully detects an anomaly with just two clusters.
+The two failure cases are a 1\% slowdown of a node, and when 99\% of the log is not useful information and information is outputted at random times.
+We assume these cases are rare, and are satisfied with successfully clustering a 5\% slowdown and 50\% junk log data.
+
+\begin{figure}[t]
     \centering
     \includegraphics[width=1.0\columnwidth]{images/anomaly.pdf}
-    \caption{Anomaly detection}
+    \caption{Results for anomaly detection via clustering.
+The x axis shows different simulation evaluations; the yaxis is the number of clusters required until the anomalous node is in its own cluster.
+Two clusters is the best-case scenario.
+See Section~\ref{subsec:results:anomaly} for a description of each of the simulation numbers.}
     \label{fig:anomaly}
@@
 
-If we were able to find a ground-truth log file with labeled anomalies, we could compare our error
-rates to those presented in papers by Wei Xu et al. Unfortunately, we couldn't find any publicly
-accessible log database.
+We would like to eventually compare our algorithm with the results of Xu and collaborators \cite{Xu09}.
+Unfortunately, all of their results used private proprietary logs, so we could not do a direct comparison.
 
-\subsection{Timing Information}
-We ran on 500,000 messages per node and on average 30MB of data were sent to the HTTP Server per
-node. Visualization took on the client side so the computation cost of visualization doesn't slow
-down the server. In Figure~\ref{fig:timings} you can see basic timing information of our system.
+\subsection{Timing}
+\label{subsec:results:timing}
+Our log files had over 500,000 messages per node.
+After compression, 30MB of data were sent to the master server per node.
+The visualization was served to the client via an HTTP server, and all javascript visualizations were computed client-side, so the only cost of visualization is generating and transferring the data.
 
-\begin{figure}[p]
+Figure~\ref{fig:timings} shows basic timing information of our system.
+From this, it is clear that the per-node cost is small: less than two seconds to compute features per-node.
+This data is then passed on to the master node, which preprocesses everything and prepares to send it to the client via HTTP.
+This preprocessing is the bottleneck in our unoptimized implementation.
+Finally, when the user changes the parameters with the web interface, recomputing the clusters takes an average of $.2$ seconds.
+
+\begin{figure}[t]
     \centering
     \includegraphics[width=1.0\columnwidth]{images/timing.pdf}
-    \caption{Timing information}
+    \caption{Timing information.
+The first column is the per-node log parsing and feature extraction.
+The second column is the work the master server must do.
+The last column is the time a user must wait until an updated visualization is displayedafter changing the configuration.
+See Section \ref{subsec:results:timing} for more details.}
     \label{fig:timings}
@@
 \section*{Acknowledgements}
-The authors appreciate Ling Huang for his helpful advices and Sean Arietta for providing the log
-data gathered from his research project.
+The authors would like to thank Ling Huang for helpful discussions and Sean Arietta for providing us with the log data gathered from his research project.
 

### 2013-12-17 duplicate_nodes.py
Add duplicate_nodes.py file

@@ -1,0 +1,22 @@
+from shutil import copytree, ignore_patterns
+import glob
+import os
+import sys
+
+
+if __name__ == '__main__':
+    data_dir = './parsedData/'
+    use_symlink = True
+
+    orig_nodes = os.listdir(data_dir)
+    orig_nodes = [os.path.basename(i) for i in glob.glob(os.path.join(data_dir, '1*'))]
+
+    for dup_cnt in range(100):
+        for orig_node in orig_nodes:
+            src = os.path.join(data_dir, orig_node)
+            dst = os.path.join(data_dir, 'd%s_%04d' % (orig_node, dup_cnt))
+            if use_symlink:
+                src = os.path.relpath(src, data_dir)
+                os.symlink(src, dst)
+            else:
+                copytree(src, dst)

### 2013-12-18 paper/paper.tex
Updated timing results in paper; prettier figs

@@
 The user can then interactively search the log, zoom in on a certain time window, and choose the number of clusters.
-%This means that at the most ``zoomed out'' state (no time window specified, no search patters, and no clustering), the raw log data is displayed.
 Our simulations show that our clustering is effective and our system is fast:
 the clustering successfully isolates anomalously-performing nodes in a variety of situations;
-the visualization can interactively visualize hundreds of millions of log messages across a hundred nodes in real-time.
-TODO: Get the numbers from this simulation.
+the visualization can interactively visualize hundreds of millions of log messages across a hundred nodes at interactive rates.
 
@@
 Their method does not seem to be, nor claim to be, interactive.
-We find their clustering method and features to be complementary to ours, and it would fit cleanly in our interactive system.
+Finally, Saganowski and collaborators present a statistical method for preprocessing features that will be used in anomaly detection \cite{preprocessor}.
+We find that these two methods for clustering, feature extraction, and feature preprocessing to be complementary to ours, and it would fit cleanly in our interactive system.
 
@@
 This preprocessing is the bottleneck in our unoptimized implementation.
-Finally, when the user changes the parameters with the web interface, recomputing the clusters takes an average of $.2$ seconds.
+Finally, when the user changes the parameters with the web interface, the k-means clustering is recomputed at interactive rates.
+
+When the number of nodes increases to 100, and the number of log messages to 50 million, we are still able to run the clustering in 2.1 seconds on a single thread.
+We have not found it necessary to use distributed or approximate k-means clustering, but if the number of nodes increases, it will be worthwhile to use a sublinear k-means algorithm.
 

### 2013-12-18 paper/paper.tex
Fix typos

@@
 
-With a large enough cluster, this will require focusing the user’s attention on “interesting” parts
+With a large enough cluster, this will require focusing the user's attention on “interesting” parts
 of the system state. This will require detecting anomalies, which in turn requires a
@@
 Munin, based on RRDtool, is similar to Ganglia: after installation on each node, it automatically presents a visualization of the system status~\cite{Oetiker99}.
-As is evident from Figure~\ref{fig:rrdtool}, the visualizations is extremely dense and illegible as the number of nodes increase.
+As is evident from Figure~\ref{fig:rrdtool}, the visualization is extremely dense and illegible as the number of nodes increases.
 
@@
     \item Compute per-node features
-    \item Perform distributed k-means clustering
+    \item Perform $k$-means clustering
     \item Display interactive visualization
@@
 \subsection{Clustering}
-We perform k-means clustering in the feature vector space~\cite{kmeans,Lloyd82}.
+We perform $k$-means clustering in the feature vector space~\cite{kmeans,Lloyd82}.
 With a small number of nodes (on the order of tens), the clustering can be performed on a single machine quickly.
-On more nodes, a distributed $k$-means algorithm is required, which can either be exact~\cite{jin2006fast} or approximate and slower~\cite{kanungo2002local}.
+On more nodes, a distributed $k$-means algorithm is required, which can either be exact and slower~\cite{jin2006fast} or approximate~\cite{kanungo2002local}.
 Since the clustering will occur each time the user updates the parameters of the visualization, it is essential that it is fast.
@@
 Once the desired number of centroids are found, we compute the node which is closest to the centroid in a least-squares sense and display only that node.
-Therefore, thhe visualization master server need only receive $O(k)$ data to pass on to the user, linear with the number of centroids, instead of with the total number of nodes.
+Therefore, the visualization master server need only receive $O(k)$ data to pass on to the user, linear with the number of centroids, instead of with the total number of nodes.
 
@@
 \label{subsec:results:anomaly}
-The metric of success for anomaly detection is as follow: assuming only one node is anomalous, how many clusters are required until that node is in its own cluster?
+The metric of success for anomaly detection is as follows: assuming only one node is anomalous, how many clusters are required until that node is in its own cluster?
 The best-case scenario is two clusters: one for each of the $n-1$ well-performing nodes, and another for the anomalous node.
@@
 
-Only one out of every hundred messages is real. One node is still 5\% slower. Can the anomalous node be detected? Note that the filesize here is bloated to 100 times the actual log data.
+Only one out of every hundred messages is real. One node is still 5\% slower. Can the anomalous node be detected? Note that the file size here is bloated to 100 times the actual log data.
 \end{enumerate}
@@
 We would like to eventually compare our algorithm with the results of Xu and collaborators \cite{Xu09}.
-Unfortunately, all of their results used private proprietary logs, so we could not do a direct comparison.
+Unfortunately, all of their results used private and proprietary logs, so we could not do a direct comparison.
 
@@
 From this, it is clear that the per-node cost is small: less than two seconds to compute features per-node.
-This data is then passed on to the master node, which preprocesses everything and prepares to send it to the client via HTTP.
 This preprocessing is the bottleneck in our unoptimized implementation.
-Finally, when the user changes the parameters with the web interface, the k-means clustering is recomputed at interactive rates.
+After the distributed preprocessing this data is then passed on to the master node, which prepares to send it to the client via HTTP.
+Finally, when the user changes the parameters with the web interface, the $k$-means clustering is recomputed at interactive rates.
 
 When the number of nodes increases to 100, and the number of log messages to 50 million, we are still able to run the clustering in 2.1 seconds on a single thread.
-We have not found it necessary to use distributed or approximate k-means clustering, but if the number of nodes increases, it will be worthwhile to use a sublinear k-means algorithm.
+We have not found it necessary to use distributed or approximate $k$-means clustering, but if the number of nodes increases, it will be worthwhile to use a sublinear $k$-means algorithm.
 

### 2013-12-18 paper/paper.tex
Updated discussion/conclusion

@@
 
-\author{Woojong Koh and Armin Samii\thanks{e-mail:\{wjkoh,samii\}@berkeley.edu}\\University of California, Berkeley}
+\author{Armin Samii and Woojong Koh\thanks{e-mail:\{samii,wjkoh\}@berkeley.edu}\\University of California, Berkeley}
 \pdfauthor{Armin Samii and Woojong Koh}
@@
 We assume that all logs are generated using the Google Logging Library (GLOG)\footnote{\url{https://code.google.com/p/google-glog/}} in order to know which source line generated which log message.
-This assumption can be dropped using the mining algorithm of Xu and collaborators~\cite{Xu:2009:OSP:1674659.1677125}.
+This assumption can be dropped using the mining algorithm of Xu and collaborators~\cite{XuOSP}.
 
@@
 
-\section{Discussion}
-% What has the audience learned about visualization from your work?
+\section{A Note on the Data}
 
-The log data we used in this project came from the system which looks through terabytes of Google
-Street View data and uses them for various machine learning tasks. It uses about 80 processes, each
-process with its own log file, and spread over around a dozen machines. There are regular failures
-which are currently cumbersome to find and debug.
+Our goal for this project was to assist in analyzing a specific source of log files, which was a machine learning algorithm on terabytes of data.
+The machine learning algorithm used about 80 processes, each with its own log file, and spread over around a dozen machines.
+There are regular failures which are currently cumbersome to find and debug.
+At the same time, we kept our system general enough to work on other types of log files, particularly one with many more than a dozen nodes.
 
-We wanted to create a responsive system which visualizes the state of the system in real-time, is
-robust to errors, and can display an arbitrary number of nodes. Not all nodes will be displayable on
-a single screen, so the system had to be able to determine areas of interest (e.g. failed nodes,
-slow nodes, or completed tasks) in the visualization and direct the user’s attention there. Our
-anomaly detection built on Wei Xu's in several ways: (1) their work does not use timestamps, which
-we believe are vital to understanding the system’s state; (2) their work intentionally ignores the
-ordering of log messages, and instead pays attention only to the number of occurrences of a message
-, we hypothesize order may matter; (3) most importantly, their visualizations are more primitive. We
-believe the idea of near-infinite zoom (where the most detailed layer essentially shows the
-information in the actual log, and as you zoom out, more and more information is aggregated) will
-allow a user to explore the system state rapidly.
+For this particular application, we believe our system can immensely simplify the analysis of the log data.
+Diverging residuals, crashed nodes, and network failures are the common failure cases, and our system can capture all of these.
+While other log analysis research often focuses on analysis of a \emph{system} \cite{XuOSP,Xu09}, we also focus on analysis of an \emph{algorithm} which periodically logs its progress.
+The failures are then in the variables, not the existence of the message.
 
-What we want to point out in this paper is that along with our project there is a direction of
-visualization research that can ultimately help system administrators to monitor and analyze the
-status of a large distributed system. We think this will become more and more important as the trend
-of technology emphasizes on parallelism and distributed system such as cloud computing.
+\section{Conclusions and Future Work}
 
-\section{Future Work}
-% A description of how your system could be extended.}
+We have presented a system for visualizing arbitrary log files.
+The log files are automatically and distributedly parsed, features are computed per node, and the results are visualized in an easy-to-use web interface.
+We cluster similarly-performing nodes into a single visualization to abstract away redundant data, if the user desires.
 
-We believe automatically determining plot type would be a good improvement for our system because
-data extracted from raw log messages can be arbitrary and it is hard to be predetermined by an user.
-Also, we want to extend our system to online and real-time analysis and implement more intelligent
-features preprocessing. Lastly, it would be great if our system can cluster logs based on more
-elaborate features.
+The visualization has a spectrum of parameters.
+On one end of this spectrum, the raw log data is shown; on the other end, a single graph shows the most representative state of the system.
+The importance of this point is \emph{trustworthiness} of the visualization: no data is hidden from the user.
+
+Our visualization is still extremely primitive.
+Users may have no intuition as to why two clusters are separated.
+Displaying a cluster confidence may assist in telling the user when two clusters are not very different.
+On the other hand, if two clusters should be different (such as in the case of a 1\% slowdown of a node), we would want to tell the user more information about why they were clustered differently.
+As is, it would be very hard for a user to notice a 1\% slowdown in a node.
+
+We would like to compute more elaborate features for the clustering, including higher-order statistics.
+However, as the feature complexity increases, the interactivity decreases.
+Finding a middle-ground between interactivity and complexity is an open problem which we are not sure how to address.
+
+Finally, we would need a more thorough evaluation of our visualization.
+Some questions we have are:
+\begin{enumerate}
+    \item is a scatterplot the best visualization choice?
+    \item are we able to successfully segregate other types of anomalies in different clusters?
+    \item can multiple messages be aggregated into the same plot?
+\end{enumerate}
+
+There is a lot to be desired before our system is usable in the wild.
+We believe we have taken a major step in the right direction by providing the user with tools to explore the raw log data, rather than trying to predict what they would like to know.
 
 \section*{Acknowledgements}
-The authors would like to thank Ling Huang for helpful discussions and Sean Arietta for providing us with the log data gathered from his research project.
+The authors would like to thank Ling Huang for helpful discussions and insight into the structure of log files,
+Sean Arietta for providing us with the log data gathered from his research project,
+and John Kubiatowicz and Anthony Joseph for their guidance throughout this project.
+Adrien Truielle at Carnegie Mellon University provided the desktop machine on which we ran our experiments.
 

### 2013-12-18 paper/paper.tex
Updated conclusion

@@
 The log files are automatically and distributedly parsed, features are computed per node, and the results are visualized in an easy-to-use web interface.
-We cluster similarly-performing nodes into a single visualization to abstract away redundant data, if the user desires.
+We cluster similarly-performing nodes into a single visualization to abstract away redundant data.
 
@@
 On one end of this spectrum, the raw log data is shown; on the other end, a single graph shows the most representative state of the system.
-The importance of this point is \emph{trustworthiness} of the visualization: no data is hidden from the user.
+This is important to maintain \emph{trustworthiness} of the visualization: no data is hidden from the user.
 
@@
 Displaying a cluster confidence may assist in telling the user when two clusters are not very different.
-On the other hand, if two clusters should be different (such as in the case of a 1\% slowdown of a node), we would want to tell the user more information about why they were clustered differently.
+On the other hand, if two clusters should be different, we would want to provide an explanation to the user similar to the decision tree visualization of Xu and collaborators \shortcite{Xu09}.
 As is, it would be very hard for a user to notice a 1\% slowdown in a node.
 
-We would like to compute more elaborate features for the clustering, including higher-order statistics.
+We would like to compute more representative features for the clustering, including higher-order statistics.
 However, as the feature complexity increases, the interactivity decreases.
-Finding a middle-ground between interactivity and complexity is an open problem which we are not sure how to address.
+Finding a middle-ground between interactivity and complexity is an difficult problem: we must be able to compute all of the features without touching all of the data.
 
@@
     \item can multiple messages be aggregated into the same plot?
+    \item what additional information would assist the human in understanding the visualization?
 \end{enumerate}
